\documentclass[12pt]{article} 

\usepackage{url,enumerate}

\pagestyle{empty} \textheight 25.0cm \textwidth 18.0cm
\oddsidemargin -1.0cm \evensidemargin -1.0cm

\title{\vspace{-2cm}CHEN10072 Engineering Mathematics 2: Overview 2017/18}
\date{\vspace{-1cm}}
\author{}
\begin{document}


\maketitle

\begin{enumerate}
\item {\bf Matrices. Week 1 (Lectures 1-2) } Matrices and algebraic properties. Definition of determinant and properties. Identity, diagonal, transpose, upper/lower triangular and inverse matrices.
\item {\bf Systems of equations. Weeks 2-3 (Lectures 3-6)} Solving systems of linear equations through Gaussian Elimination. Solving systems of nonlinear equations using the Newton-Raphson method.
\item {\bf Analytic ordinary differential equations. Week 4 (Lectures 7-8)} First order ordinary differential equations (ODEs). Analytical solution by separation of variables and by the integrating factor.
\item {\bf Laplace Transforms. Weeks 5-6 (Lectures 9-12)} Laplace transforms. Transforms and inverse transforms of functions. Transforms of derivatives. Application to solution of ordinary differential equations.
\item {\bf Numerical ordinary differential equations. Weeks 7-8 (Lectures 13-16)}  Numerical solution of ODEs and systems of ODEs using Euler and Runge-Kutta Methods.
\end{enumerate}


\begin{thebibliography}{9}
\bibitem{1} Stroud KA and Booth DJ. \textit{Engineering Mathematics}. $5^{th}$ edition. Palgrave, 2001.
\bibitem{2} Chapra SC and Canale RP. \textit{Numerical methods for engineering}. $7^{th}$ edition. McGraw-Hill, 2006.
\bibitem{3} Quarteroni A, Saleria F and Gervasio P. \textit{Scientific Computing with MATLAB and Octave}. $4^{th}$ edition. Springer, 2010.
\bibitem{4} HELM: Helping Engineers Learn Mathematics \url{http://www.maths.manchester.ac.uk/study/undergraduate/information-for-current-students/service-teaching/resources/}
\end{thebibliography}

\end{document}